\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Assessing income inequality with tax data - Switzerland from 1917 to 2010}

### 2014-07-03 paper/subdocuments/introduction.Rnw
viele mini fixes aus gemeinsamer session

% 3.Abschnitt: What do we know about Switzerland
% Offical Data Collections on development of inequality
As we focus our paper on the case of Switzerland, it is important to embed our work in the context of given publications concerning inequality in income. What is known about Switzerland so far? Looking for official data, three main sources has to be mentioned, which can be considered as de facto official data sources: EU-SILC, HBS and LIS-data. Figure XY shows the results stemming from this three sources while looking at Gini of equivalised disposable income. Up to the day, EU-SILC or Statistics on Income and Living Conditions is the main source used for policy monitoring at EU-level. The main focus of EU-SILC is to collect data on a common ``framework'' to ensure comparability among EU-countries and countries living around or within the EU. As a Non-EU member Switzerland implemented the instrument not from the beginning (2004) but as from 2007. Therefore this times-series doesn't cover time points before 2007. As graph XY shows, following the results from SILC income inequality decreased from 2007 to 2013.\footnote{Data shown in the graph was downloaded from the Eurostat Metadata-portal \url{http://epp.eurostat.ec.europa.eu/cache/ITY_SDDS/EN/ilc_esms.htm} last accessed 21.Mai 2014.} The second important source concerning the distribution of income is the Household Budget Survey (HBS). The main focus of this survey lays in providing detailed data on household budgets. This allows researcher to look at different income concepts like income before and after public transfers. Since 2000 the survey has been conducted on a continuous basis, which allows to look at a consistent time series from 2000 to 2011. As it can be seen from graph XY the trend is rather stable.\footnote{figures shown in the graph were calculated out of the original datasets, which were kindly provided by Swiss Federal Statistical Office.} Both time-series (SILC and HBS) cover a relatively short time period. A longer period is covered in the LIS-Data-set (1982-2004). Data-provider for the LIS Data is the Swiss Federal Statistical Office too. In contrast the the aforementioned surveys the LIS-data is harmonized out of three surveys: Swiss Income and Wealth Survey (1982), Swiss Poverty Survey (1992) and the Income and Consumption survey (2000, 2002, 2004).  All in all the LIS dataset contains the longest time series on inequality for Switzerland. Analyzing this data \citet{gornick_income_2013} found for Switzerland a quite substantially decreases in income inequality, contradictory to the development in most other western countries. This result is supported by \citet{grabka_evolution_2012} analyzing the Swiss Household Panel (2000-2009).\footnote{A further official database for income distribution is the OECD-Database. It includes measures from Income and Consumption survey as well. Additional data for 2008 is available from EU-Survey of Income and Living Conditions (EU-SILC). But this change in survey is considered as a strict break. Comparison before and since 2008 is not recommended (OECD 2012:315). For the sake of completeness the database construtcted by \citet{deininger_new_1996} and the World Income Inequality Database (WIID) have to be mentioned. Both datasets do not contain figures  for recent years for Switzerland. A further important database on inequality is the GINI database which has been derived from the GINI Country Reports \citep{nolan_changing_2014}. But this dataset doesn't cover Switzerland.}  \\

<<Swiss inequality Plot>>=
library(ggplot2)
library(foreign)
library(reshape, quietly=TRUE, warn.conflicts = FALSE)
number_ticks <- function(n) {function(limits) pretty(limits, n)}
swissGini<-read.csv("../../data/swissGini.csv",header=TRUE,sep=";")
swissGini_long<-melt(swissGini,id="Year")
ggplot(data=swissGini_long,
  aes(y=value,x=Year),shape=variable)+
  ylab("Gini of equivalised disposable income") +
  scale_y_continuous(limits=c(0.18,0.4)) +
  scale_x_continuous(limits=c(1982,2012),breaks=number_ticks(10)) +
  geom_point(aes(shape=variable),size=3)+
  geom_line(data=swissGini_long[!is.na(swissGini_long$value),],aes(linetype=variable))+
  theme_bw()
@

@@

% 4.Abschnitt: What do we know about Switzerland
% Studies with tax data
Whereas the aforementioned publications focused on disposable household income from survey data, the revival of tax-data-inequality studies lead to fruitful insights for Switzerland as well. \citet{dell_income_2007} used tax data from the Federal Tax Administration to assess the development of concentration of the highest incomes and wealth (top-shares). In contrast to most other examined countries, Switzerland did not experience a reduction in income and wealth concentration from the pre-First World War period to the decades following the second World War (up to 1996). Using the same approach \citet{foellmi_volatile_2013} expand the Dell et al. time series to 2008 finding that the share of top income has risen, the top 0.01\% share even doubled in the last observed 20 years. A result which opposes the outcome of official data published by the Swiss Federal Statistical Office.\footnote{There are other studies on Switzerland covering different periods but not the recent years. \citet{fluckiger_egalite_2002} and also \citet{jeitziner_regionale_2007, jeitziner_regionale_2009} report constant inequality from 1960-1996 respectively from 1995 to 2003. Covering a similar time period \citet{bauer_verteilung_1994} and \citet{bolzani_distribution_2001} found decreasing inequality. On the other hand \citet{buchmann_zur_1995} and \citet{ernst_inequality_2000} found an increase in the 1980s).} \\

% 5. Abschnitt: Zwischenbilanz und Herausforderungen
Divergence can be explained with several factors. First of all, different data sources were used. The official data providers trust on survey data, whereas the later mentioned publications use tax data. It is assumed that the coverage of top incomes is better in tax data than it is in survey data (non-respondent bias), which is a crucial issue concerning inequality. On the other hand the focus on top income neglects other changes in the distribution of income as it is not possible to see, whether newer concerns like the ``hollowing of the middle class'' occurred in Switzerland or not, which leads to the second point. Different measure of inequality hampers the comparability. Third, different income concepts and different units of analysis were used. As it is shown by \citet{modetta_einkommensungleichheit_2012} income distribution is strongly affected by governmental redistribution, reducing inequality substantially. With the focus on tax data the change in institutional settings is not covered. Also neglected is the household structure, whereas it is unclear how inequality is affected whether one looks at household income or at income of tax units. It can be assumed, that inequality corresponding to different concepts react differently on demographic change (change in household structure).

@@

Up to the day, Switzerland can be situated according to the actual level of income inequality in western societies as there is a huge effort to collect data which can be harmonized to comparable measures (see Luxembourg income study, EU-SILC). However, it is unclear how the bias through non-response affects the overall measure of inequality. For the US \citet{atkinson_top_2009} estimate that CPS survey data fail to capture about half of the overall increase in inequality measured by the Gini coefficient, a result confirmed by \citet{alvaredo_note_2011}. Likewise a long and consistent time-series allowing to identify and explore development patterns on every point of the distribution (not only the top-shares) is missing. Building on recent developments in the field of inequality research, we assess the suitability of the publicly accessible tax data to report inequality and its changes over time. First of all this includes a discussion of the accessible measures in context of a reflection about the state of the art conceptualization of economic resources as an indicator for economic well-being. Second, we summarize and apply tax data specific techniques to construct suitable measures of inequality. We expand the given set by an in depth discussion of a newly applied step to handle the incomplete coverage in tax data statistics. Third, we compare our results to results from a relevant Inequality-Survey in Switzerland (HBS) to assess the bias through data source.

>>>>

% 3.Abschnitt: What do we know about Switzerland
% Offical Data Collections on development of inequality
As we focus our paper on the case of Switzerland, it is important to embed our work in the context of given publications concerning inequality in income. What is known about Switzerland so far? Looking for official data, three main sources has to be mentioned, which can be considered as de facto official data sources: EU-SILC, HBS and LIS-data. Figure \ref{fig:SwissinequalityPlot} shows the results stemming from this three sources while looking at Gini of equivalised disposable income. Up to the day, EU-SILC or Statistics on Income and Living Conditions is the main source used for policy monitoring at EU-level. The main focus of EU-SILC is to collect data on a common ``framework'' to ensure comparability among EU-countries and countries living around or within the EU. As a Non-EU member Switzerland implemented the instrument not from the beginning (2004) but as from 2007. Therefore this times-series doesn't cover time points before 2007. As graph XY shows, following the results from SILC income inequality decreased from 2007 to 2013.\footnote{Data shown in the graph was downloaded from the Eurostat Metadata-portal \url{http://epp.eurostat.ec.europa.eu/cache/ITY_SDDS/EN/ilc_esms.htm} last accessed 21.Mai 2014.} The second important source concerning the distribution of income is the Household Budget Survey (HBS). The main focus of this survey lays in providing detailed data on household budgets. This allows researcher to look at different income concepts like income before and after public transfers. Since 2000 the survey has been conducted on a continuous basis, which allows to look at a consistent time series from 2000 to 2011. As it can be seen from figure \ref{fig:SwissinequalityPlot} the trend is rather stable.\footnote{figures shown in the graph were calculated out of the original datasets, which were kindly provided by Swiss Federal Statistical Office.} Both time-series (SILC and HBS) cover a relatively short time period. A longer period is covered in the LIS-Data-set (1982-2004). Data-provider for the LIS Data is the Swiss Federal Statistical Office too. In contrast to the aforementioned surveys the LIS-data is harmonized out of three surveys: Swiss Income and Wealth Survey (1982), Swiss Poverty Survey (1992) and the Income and Consumption survey (2000, 2002, 2004).  All in all the LIS dataset contains the longest time series on inequality for Switzerland. Analyzing this data \citet{gornick_income_2013} found for Switzerland a quite substantially decreases in income inequality, contradictory to the development in most other western countries. This result is supported by \citet{grabka_evolution_2012} analyzing the Swiss Household Panel (2000-2009).\footnote{A further official database for income distribution is the OECD-Database. It includes measures from Income and Consumption survey as well. Additional data for 2008 is available from EU-Survey of Income and Living Conditions (EU-SILC). But this change in survey is considered as a strict break. Comparison before and since 2008 is not recommended (OECD 2012:315). For the sake of completeness the database construtcted by \citet{deininger_new_1996} and the World Income Inequality Database (WIID) have to be mentioned. Both datasets do not contain figures  for recent years for Switzerland. A further important database on inequality is the GINI database which has been derived from the GINI Country Reports \citep{nolan_changing_2014}. But this dataset doesn't cover Switzerland.}  \\

<<SwissinequalityPlot, fig.cap='Trends in income inequality',fig.pos='H'>>=
library(ggplot2)
library(foreign)
library(reshape, quietly=TRUE, warn.conflicts = FALSE)
number_ticks <- function(n) {function(limits) pretty(limits, n)}
swissGini<-read.csv("../../data/swissGini.csv",header=TRUE,sep=";")
swissGini_long<-melt(swissGini,id="Year")
swissGini_long$Source<-swissGini_long$variable
ggplot(data=swissGini_long,
  aes(y=value,x=Year),shape=Source)+
  ylab("Gini of equivalised disposable income") +
  scale_y_continuous(limits=c(0.18,0.4)) +
  scale_x_continuous(limits=c(1982,2012),breaks=number_ticks(10)) +
  geom_point(aes(shape=Source),size=3)+
  geom_line(data=swissGini_long[!is.na(swissGini_long$value),],aes(linetype=Source))+
  theme_bw()
@

@@

% 4.Abschnitt: What do we know about Switzerland
% Studies with tax data
Whereas the aforementioned publications focused on disposable household income from survey data, the revival of tax-data-inequality studies lead to fruitful insights for Switzerland as well. \citet{dell_income_2007} used tax data from the Federal Tax Administration to assess the concentration of the highest incomes and wealth (top-shares) over time. In contrast to most other examined countries, Switzerland did not experience a reduction in income and wealth concentration from the pre-First World War period to the decades following the second World War (up to 1996). Using the same approach \citet{foellmi_volatile_2013} expand the Dell et al. time series to 2008 finding that the share of top income has risen, the top 0.01\% share even doubled in the last observed 20 years. A result which opposes the outcome of official data published by the Swiss Federal Statistical Office.\footnote{There are other studies on Switzerland covering different periods but not the recent years. \citet{fluckiger_egalite_2002} and also \citet{jeitziner_regionale_2007, jeitziner_regionale_2009} report constant inequality from 1960-1996 respectively from 1995 to 2003. Covering a similar time period \citet{bauer_verteilung_1994} and \citet{bolzani_distribution_2001} found decreasing inequality. On the other hand \citet{buchmann_zur_1995} and \citet{ernst_inequality_2000} found an increase in the 1980s).} \\

% 5. Abschnitt: Zwischenbilanz und Herausforderungen
Divergence can be explained with several factors. First of all, different data sources were used. The official data providers trust on survey data, whereas the later mentioned publications use tax data. It is assumed that the coverage of top incomes is better in tax data than it is in survey data (non-respondent bias), which is a crucial issue concerning inequality. On the other hand the focus on top income neglects other changes in the distribution of income as it is not possible to see, whether newer concerns like the ``hollowing of the middle class'' occurred in Switzerland or not, which leads to the second point. Different measure of inequality hampers the comparability. Third, different income concepts and different units of analysis were used. As it is shown by \citet{modetta_einkommensungleichheit_2012} the income distribution is strongly affected by governmental redistribution, reducing inequality substantially. With the focus on tax data the change in institutional settings is not covered. Also neglected is the household structure, whereas it is unclear how inequality is affected whether one looks at household income or at income of tax units. It can be assumed, that inequality corresponding to different concepts react differently on demographic change (change in household structure).

@@

Up to the day, Switzerland can be situated according to the actual level of income inequality in western societies as there is a huge effort to collect data which can be harmonized to comparable measures (see Luxembourg income study, EU-SILC). However, it is unclear how the bias through non-response affects the overall measure of inequality. For the US \citet{atkinson_top_2009} estimate that CPS survey data fail to capture about half of the overall increase in inequality measured by the Gini coefficient, a result confirmed by \citet{alvaredo_note_2011}. Likewise a long and consistent time-series allowing to identify and explore development patterns on every point of the distribution (not only the top-shares) is missing. Building on recent developments in the field of inequality research, we assess the suitability of the publicly accessible tax data to report inequality and its changes over time. First of all this includes a discussion of the accessible measures in context of a reflection about the state of the art conceptualization of economic resources as an indicator for economic well-being (see section \textit{II. Data and measurement concepts}). Second, we examine the role of methodological decisions along with an in-depth distributional analysis (see sections \textit{III. Methods} and \textit{IV. Results}). Against the background of all methodological issues we discuss the long-term inequality development in Switzerland -- as well as more recent periods in detail -- using all relevant data.

### 2014-07-03 paper/subdocuments/methods.Rnw
viele mini fixes aus gemeinsamer session

@@

To fulfill the above described task, we use two techniques. To assess the development over time, we calculate in general gini-coefficients for all possible time points, allowing us the make time trends visible and to compare the result out of the FTA-tax data to other publications. For selected periods we expand the analysis with relative distribution methods, which allows an in-depth analysis of distributional differences and therefore catches the shortcomings of gini-coefficients. \\

To calculate Gini-coefficients we follow common definitions, like it is described for example in \citep{jann_einfuhrung_2005}. Because the relative distribution methods are not that widely know, we provide a short introduction of the important concepts in this section. We additional expand the method section with an extended analysis about the importance of non-taxed, because the issue of non-taxed is a especially thorny one.

@@

The RD-framework is based on the concept of a ``relative distribution'', a transformation of the data from two distributions into a single distribution that contains all of the information necessary for scale-invariant comparison. This allows to make distribution differences ``visual'' in an elegant way and it is also a base for summary statistics, which are more sensitive to detailed theoretical hypotheses in contrast to other measures like the gini-coefficient or top income-shares, which inform either about the whole population or only about the top part of the distribution.\\

@@

R is obtained from $Y$ by transforming it by the CDF for $Y_{0}$, $F_{0}$. $R$ therefore measures the relative rank of $Y$ compared to $Y_{0}$.

@@

We can calculate the Probability Density Function $g(r)$ of R, where $r$ represents the proportion of values and $F_{0}^{-1}(r)$ is the inverse cumulative distribution function, also called the quantile function. $g(r)$ can be interpreted as a density ratio, which is defined as the ratio of these two quantities evaluated at every percentile of the reference distribution [0,1]. With a complete overlap of both distributions the probability density function of the R is 1 at every point of the PDF. On the other hand, values higher than 1 represent higher probabilities in the comparison distribution than in the references distribution at this specific point and values lower than 1 respectively represent lower probabilities. It is a proper PDF in the sense that it integrates to 1 over the unit interval.\\

@@

% R based Summary Meausres
Distributional polarization is of particular interest in the study of inequality. However, common inequality indicators (for example Gini or Theil’s index) are not designed to distinguish between growth in the upper and lower tails. Even if the measures register increasing inequality over time, one cannot distinguish a polarization of the distribution (increases in both tails) from upgrading (increases in the upper tail) or downgrading (increases in lower tail). The polarization index developed by \citep{Handock and Morris 1999} adresses this issue, because it is decomposable to distinguish differences in the upper and lower tails.  Because it is based on the relative distribution it provides a simple link between what is observed in the graphical display and what is measured by the numerical summary. \\

The median relative polarization index (MRP) is defined as the mean absolute deviation around the median of the location-adjusted relative distribution, scaled to produce an index that varies between -1 and 1. Given the scaling, a value of zero represents no differences in distributional shape; positive values represent more polarization (increases in the tails of the distribution); and negative values represent less polarization (convergence towards the center of the distribution). The measure catches only differences in distributional shape (not location). And it has several interesting features. MRP can be interpreted in terms of a proportional shift of mass in the distribution from more central to less central values. A value of 0.1, for example, is equivalent to a 10\% population shift from the center of the distribution to the upper and lower quartiles and the MRP is decomposable along the scale of $y$. This makes it possible to compare the contribution of each section of the distribution to the overall polarization. A natural decomposition is the contributions made by components above and below the median (of $g(r)$).

@@

<<zero_descriptives, fig.cap='',fig.pos='H'>>=
library(ggplot2)
ggplot(df[df$steuerperiode>=1995&df$kanton=="CH",], aes(x=steuerperiode,y=zeros))+geom_line()+theme_bw()+xlab("tax period")+ylab("share of non-taxed")
#ggplot(df[df$steuerperiode>=1995,], aes(x=steuerperiode,y=zeros))+geom_line()+facet_wrap(~kanton)+geom_line(aes(y=G_steink),color="blue")+theme_bw()
@

@@

By adding the zeros back as a separate group we need to make an assumption about the income distribution within that group. We know these cases had incomes between zero and the threshold (usually around 15.000 CHF for the period 1995-2010). In figure \ref{fig:with_without_zeros_Gini} we show Gini coefficients calculated under the assumption of the zeros being zero. This is problematic for two reasons: first we know these units might earn more (up to the threshold) so Ginis calculated on this assumptions will be too high. Second we assume the zero group to be perfectly homogenous which on the contrary has a lowering effect on the Gini coefficient. Therefore we must at least assume a distribution for the zeros that has some variance and ranges from zero to the threshold. We decided to use the most simple approach fulfilling both criteria which is to interpolate a uniform distribution between zero and each tax periods threshold.

The second approach investigates whether the FTA time series is suitable to analyze trends of inequality even if the inequality measures are biased. We can do this by regressing the inequality measure (here the Gini coefficient) on the share of zeros. Simply spoken we test if cutting a larger (left) tail of the distribution changes the shape of the distribution. We include Gini coefficient for all cantons in our analyses to improve the power of our statistical test\footnote{Furthermore without cantonal data (but only data for aggregate Switzerland) it is impossible to distinguish whether inequality changed due to the number of unobserved people or any other variable that changed over time)}.

<<corrected_gini, dev='bmp', fig.cap='',fig.pos='H'>>=
df$J<-factor(df$steuerperiode)
fit <- lm(G_steink~kanton+kanton:zeros+J+zeros:J,data=df)
#summary(fit)
SmoothCoefficientPlot <- function(models, modelnames = "", removeintercept = FALSE){
  # models must be a list()

>>>>

\section{Methods}

@@

To fulfill the above described task, we use two techniques. To assess the development over time, we calculate in general Gini-coefficients for all possible time points, allowing us the make time trends visible and to compare the result out of the FTA-tax data to other publications. For selected periods we expand the analysis with relative distribution methods, which allows an in-depth analysis of distributional differences and therefore catches the shortcomings of Gini-coefficients. \\

To calculate Gini-coefficients we follow common definitions, like it is described for example in \citep{jann_einfuhrung_2005}. Because the relative distribution methods are not that widely known, we provide a short introduction of the important concepts in this section. We additional expand the method section with an extended analysis about the importance of non-taxed, because the issue of non-taxed is a especially thorny one.

@@

The RD-framework is based on the concept of a ``relative distribution'', a transformation of the data from two distributions into a single distribution that contains all of the information necessary for scale-invariant comparison. This allows to make distribution differences ``visual'' in an elegant way and it is also a base for summary statistics, which are more sensitive to detailed theoretical hypotheses in contrast to other measures like the Gini-coefficient or top income-shares, which inform either about the whole population or only about the top part of the distribution.\\

@@

$R$ is obtained from $Y$ by transforming it by the CDF for $Y_{0}$, $F_{0}$. $R$ therefore measures the relative rank of $Y$ compared to $Y_{0}$.

@@

We can calculate the Probability Density Function $g(r)$ of R, where $r$ represents the proportion of values and $F_{0}^{-1}(r)$ is the inverse cumulative distribution function, also called the quantile function. $g(r)$ can be interpreted as a density ratio, which is defined as the ratio of these two quantities evaluated at every percentile of the reference distribution [0,1]. With a complete overlap of both distributions the probability density function of the $R$ is 1 at every point of the PDF. On the other hand, values higher than 1 represent higher probabilities in the comparison distribution than in the references distribution at this specific point and values lower than 1 respectively represent lower probabilities. It is a proper PDF in the sense that it integrates to 1 over the unit interval.\\

@@

% R based Summary Meausres
Distributional polarization is of particular interest in the study of inequality. However, common inequality indicators (for example Gini or Theil’s index) are not designed to distinguish between growth in the upper and lower tails. Even if the measures register increasing inequality over time, one cannot distinguish a polarization of the distribution (increases in both tails) from upgrading (increases in the upper tail) or downgrading (increases in lower tail). The polarization index developed by \citep{Handock and Morris 1999} addresses this issue, because it is decomposable to distinguish differences in the upper and lower tails.  Because it is based on the relative distribution it provides a simple link between what is observed in the graphical display and what is measured by the numerical summary. \\

The median relative polarization index (MRP) is defined as the mean absolute deviation around the median of the location-adjusted relative distribution, scaled to produce an index that varies between -1 and 1. Given the scaling, a value of zero represents no differences in distributional shape; positive values represent more polarization (increases in the tails of the distribution); and negative values represent less polarization (convergence towards the center of the distribution). The measure catches only differences in distributional shape (not location). And it has several interesting features. MRP can be interpreted in terms of a proportional shift of mass in the distribution from more central to less central values. A value of 0.1, for example, is equivalent to a 10\% population shift from the center of the distribution to the upper and lower quartiles and the MRP is decomposable along the scale of $y$. This makes it possible to compare the contribution of each section of the distribution to the overall polarization. A natural decomposition is the contributions made by components above (upper polarization index, URP) and below (lower polarization index, LRP) the median (of $g(r)$).

@@

<<zero_descriptives, fig.cap='Share of tax units below federal tax liability threshold',fig.pos='H'>>=
library(ggplot2)
ggplot(df[df$steuerperiode>=1995&df$kanton=="CH",], aes(x=steuerperiode,y=zeros))+geom_line()+theme_bw()+xlab("tax period")+ylab("share of non-taxed")
#ggplot(df[df$steuerperiode>=1995,], aes(x=steuerperiode,y=zeros))+geom_line()+facet_wrap(~kanton)+geom_line(aes(y=G_steink),color="blue")+theme_bw()
@

@@

By adding the zeros back as a separate group we need to make an assumption about the income distribution within that group. We know these cases had incomes between zero and the threshold (usually around 15.000 CHF for the period 1995-2010). In figure \ref{fig:with_without_zeros_Gini} we show Gini-coefficients calculated under the assumption of the zeros being zero. This is problematic for two reasons: first we know these units might earn more (up to the threshold) so Ginis calculated on this assumptions will be too high. Second we assume the zero group to be perfectly homogeneous which on the contrary has a lowering effect on the Gini coefficient. Therefore we must at least assume a distribution for the zeros that has some variance and ranges from zero to the threshold. We decided to use the most simple approach fulfilling both criteria which is to interpolate a uniform distribution between zero and each tax periods threshold.

The second approach investigates whether the FTA time series is suitable to analyze trends of inequality even if the inequality measures are biased. We can do this by regressing the inequality measure (here the Gini-coefficient) on the share of zeros. Simply spoken we test if cutting a larger (left) tail of the distribution changes the shape of the distribution. We include Gini-coefficient for all cantons in our analyses to improve the power of our statistical test\footnote{Furthermore without cantonal data (but only data for aggregate Switzerland) it is impossible to distinguish whether inequality changed due to the number of unobserved people or any other variable that changed over time)}.

<<corrected_gini, fig.cap='Bias variation by time and cantons',fig.pos='H'>>=
df$J<-factor(df$steuerperiode)
fit <- lm(G_steink~kanton+kanton:zeros+J+zeros:J,data=df)
#summary(fit)
SmoothCoefficientPlot <- function(models, modelnames = "", removeintercept = FALSE){
  # models must be a list()

### 2014-07-03 paper/subdocuments/results.Rnw
viele mini fixes aus gemeinsamer session

\emph{Equivalence scale}

@@

# Prop-Density Plot
density1 <- density(log(bd.2010.ind))
plot(x = (density1$x), y = density1$y, type = "l",
     xlab = "taxable income", ylab = "density"
     #axes = FALSE,
     #xlim = c(-50, 200),
     #ylim=c(0,0.0116031)
     )
title(main="all tax units vs without non-taxed",cex=0.6)
axis(side = 1)
axis(side = 2)
fig1legend <- list(x=c(6,6),y=c(0.6,0.6))
legend(fig1legend,lty=1:2, bty="n",
       legend=c("without non-taxed","all tax units"))
density2 <- density(log(bd.0.2010.ind))
lines(x = (density2$x), y = density2$y, type = "l",lty=2)

@@

gA0 <- reldist(y=bd.2010.ind, yo=bd.0.2010.ind,
               smooth=0.4, ci=FALSE,
               show="residual",
               bar=TRUE, quiet=FALSE,
               ylim=c(0.5,2), ylab="",
               discrete=FALSE,
               xlab="Proportion of 2010 Distribution (including non-taxed)")
title(main=paste("Effect of different shape"))
abline(h=1,lty=2)
@

@@

<<with_without_zeros_RelDist_mitESTV,warning=FALSE, fig.cap='',fig.pos='H'>>=

## Nicht fertig angepasst!!!
estvSteink<-(read.csv("../../data/ginis_und_perzentile_normal.csv", header=TRUE,sep="\t"))
estvSteink<-filter(estvSteink,kanton=="CH")
start <- which(names(estvSteink)=="p1")
end <- which(names(estvSteink)=="p9999")
percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)
estvReink<- data.frame(t(estvReink[1,start:end]),percentile)

names(bd.2010)[1]=c("inc")
bd.0.2010<- data.frame(t(bd.0.2010[1,start:end]),percentile)
names(bd.0.2010)[1]=c("inc")

# Prop-Density Plot
density1 <- density(log(bd.2010.ind))
plot(x = (density1$x), y = density1$y, type = "l",
     xlab = "taxable income", ylab = "density"
     #axes = FALSE,
     #xlim = c(-50, 200),
     #ylim=c(0,0.0116031)
     )
title(main="all tax units vs without non-taxed",cex=0.6)
axis(side = 1)
axis(side = 2)
fig1legend <- list(x=c(6,6),y=c(0.6,0.6))
legend(fig1legend,lty=1:2, bty="n",
       legend=c("without non-taxed","all tax units"))
density2 <- density(log(bd.0.2010.ind))
lines(x = (density2$x), y = density2$y, type = "l",lty=2)

gA0 <- reldist(y=bd.2010.ind, yo=bd.0.2010.ind,
               smooth=0.4, ci=FALSE,
               show="residual",
               bar=TRUE, quiet=FALSE,
               ylim=c(0.5,2), ylab="",
               discrete=FALSE,
               xlab="Proportion of 2010 Distribution (including non-taxed)")
title(main=paste("Effect of different shape"))
abline(h=1,lty=2)
@

<<polindexzeros2, results='asis'>>=
polTable(bd.2010.ind,bd.0.2010.ind,tablabel="polzeros2")
@

>>>>

\subsection{Statistical units -- Equivalence scale}

@@

par(mfrow=c(1,2))
# Prop-Density Plot
density1 <- density(log(bd.2010.ind))
plot(x = (density1$x), y = density1$y, type = "l",
     xlab = "taxable income", ylab = "density"
     #axes = FALSE,
     #xlim = c(-50, 200),
     #ylim=c(0,0.0116031)
     )
title(main="all tax units vs without non-taxed",cex=0.6)
axis(side = 1)
axis(side = 2)
fig1legend <- list(x=c(6,6),y=c(0.6,0.6))
legend(fig1legend,lty=1:2, bty="n",
       legend=c("without non-taxed","all tax units"))
density2 <- density(log(bd.0.2010.ind))
lines(x = (density2$x), y = density2$y, type = "l",lty=2)

@@

gA0 <- reldist(y=bd.2010.ind, yo=bd.0.2010.ind,
               smooth=0.4, ci=FALSE,
               show="residual",
               bar=TRUE, quiet=FALSE,
               ylim=c(0.5,2), ylab="",
               discrete=FALSE,
               xlab="Proportion of 2010 Distribution (including non-taxed)")
title(main=paste("Effect of different shape"))
abline(h=1,lty=2)

par(mfrow=c(1,1))
@

@@
